{\itshape Advanced Robotics} A simple path optimization method for motion planning

### 2015-10-22 paper.tex
Fix mistakes, subfigures numbers

% The following packages can be found on http:\\www.ctan.org
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{color}
\usepackage{graphicx} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{algorithm}
\usepackage{algpseudocode}  %\usepackage[noend]{algpseudocode}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{proof}
\newtheorem{property}{Property}
\usepackage[numbers]{natbib}
\algnewcommand\algorithmicinput{\textbf{input:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\algorithmicoutput{\textbf{output:}}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
\makeatletter
\makeatother
\usepackage{hyperref}

@@

\begin{abstract}
Most algorithms in probabilistic sampling-based path planning compute 
collision-free paths made of straight line segments lying in the configuration 
space. Due
to the randomness of sampling, the paths make detours that need to be optimized.
The contribution of this paper is to propose a gradient-based (GB) algorithm that
transforms a polygonal collision-free path into a shorter one, while both:
\begin{itemize}
\item requiring only collision checking, and not any time-consuming obstacle distance computation,
\item constraining only part of the configuration variables that may cause a 
collision, and not the entire configurations, and
\item reducing parasite motions that are not useful for the problem resolution.
\end{itemize}
The algorithm is simple and requires few parameter tuning. Experimental
results include navigation and manipulation tasks, e.g. an robotic arm 
manipulating throw a window and a PR2 robot working in a kitchen environment, 
and comparisons with a random shortcut optimizer.

@@

Motion planning for systems in cluttered environments has been addressed for more
than thirty years~\citep{ref-motionplan},~\citep{ref-pianomover}. Most planners
today randomly sample the system configuration space~\cite{ref-spaceconfig} in
order to find a collision-free path. The main issue using these techniques is
 that the computed path makes unnecessary detours and needs to be post-processed 
 before being executed by a virtual or real robot. Alternative
strategies exist however.
\begin{itemize}
\item Planning by path-optimization~\cite{itomp2012},~\cite{voronoiOMP} where
obstacle avoidance is handled by constraints or cost using computation of the
nearest obstacle distance. Most of these planners are using non-linear
optimization~\cite{BettsNonlinopt} under constraints.

Such planners provide close-to-optimality paths and have smaller time computation
for easy problems, but they are mostly unable to solve narrow passage issues.

@@

For this shortening purpose, random shortcut (RS) methods are
still very 
popular~\cite{Sekhavat-Svestka1998,Geraerts04clearancebased,HauserFastSmooth}. 
However, RS requires fine 
tuning of the termination condition (see 
Algorithm~\ref{algo:random-shortcut}) and is no efficient due to randomness for long 
trajectories where only a small part needs to be optimized. 
Figure~\ref{decoupled_DOF_optimization} presents another situation where RS will 
always fail to optimize the initial path, since it cannot decouple the 
robot degrees of freedom (DOF) on which the optimization occurs, contrary to our 
algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} 
that shortcut only one or a group of DOF.

\begin{figure}[H]
	\centering
	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
	\caption{Example of a path which random shortcut will never manage to 
	optimize: each shortcut tentative will provide a collision or will not 
	decrease the path length. The optimal path belongs to the $x-y$ plane 
	containing $q_{init}$ and $q_{end}$.}
	\label{decoupled_DOF_optimization}
\end{figure}

@@

\section{RELATED WORK}

CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
input. It minimizes a time invariant cost function using efficient covariant
hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
high velocities) and an obstacle avoidance term, provided by the distance to the 
nearest obstacle for each iteration of the trajectory. Calculating these nearest 
distances however is time-consuming because the distances between all pairs of 
objects must be computed at each time step along the path. To reduce the 
computation time, the method starts by building offline a map of distances that 
will be called during the optimization at the requested time. Besides, meshes 
are pre-processed into bounding spheres so that distances are computed faster 
at the cost of a geometry approximation.

@@

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints. 
~\cite{Guernane2011}, also based on random shooting, is guiding configuration-
generation with local holonimic considerations. But this method remains only locally 
optimal, and is not addressing high-DOF problems.
~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
using medial axis retractation for clearance and a PRS (applied only on random DOF) 
which can address the problem of Figure~\ref{decoupled_DOF_optimization} but is 
relatively slower than RS. Moreover, PRS is not using the information returned by 
the checker of which links are colliding to guide which DOF or group of DOF are 
relevant to shortcut, contrary to our method.

In some way, our method shares similarities with~\cite{PanSmoothSplineShort} 
since this latter method relies on collision checking and backtracks when an 
iteration is detected in collision, instead of trying to constantly satisfy 
distance constraints. Unlike our method however, interpolating configuration -which 
can be viewed as constraints- are freezing the whole robot configuration instead of 
a relevant subpart. Thus, this method cannot solve the 
Figure~\ref{decoupled_DOF_optimization} problem.

@@

A robot is defined by a kinematic chain composed of a tree of joints. We denote 
by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $1\leq i\leq 
m$ is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
is the dimension of $J_i$ in the configuration space, to the space 
of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
in the frame of its parent. In the examples shown in this paper, we consider 4 
types of joints described in Table~\ref{tab:joints}.
\begin{table}[b]
\centerline {
  \begin{tabular}{cccc}
    Name & dimension & config space & velocity\\
    \hline
    translation & 1 & $\real$ & $\real$\\
    bounded rotation & 1& $\real$ & $\real$\\
    unbounded rotation & 2 & $\Sone\subset\real^2$ & $\real$\\
    $SO(3)$ & 4 & $\Sthree\subset\real^4$ & $\real^3$
  \end{tabular}
}
\caption {Translation and rotation joint position are defined by 1 
parameter corresponding respectively to the translation along an axis and a 
rotation angle around an axis. Unbounded rotation is defined by a point on the 
unit circle of the plane: 2 parameters corresponding to the cosine and the 
sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
velocity of translation and unbounded rotation joints is the derivative of the 
configuration variable. The velocity of an unbounded rotation joint corresponds 
to the angular velocity. The velocity of a $SO(3)$ joint is defined by the 
angular velocity vector $\omega\in\real^3$.}
\label{tab:joints}
\end{table}
A configuration of the robot
$$\conf = (\underbrace{q_1,\cdots,q_{n_1}}_{J_1},\underbrace{q_{n_1+1},
\cdots,q_{n_1+n_2}}_{J_2},\cdots q_n),\ n\triangleq\sum_{i=1}^m n_i$$
is defined by the concatenation of the joint configurations. The configuration 
space of the robot is denoted by $\CS\subset\real^n$.

@@

Note that the definitions above stem from the Riemanian structure of the 
configuration space of the robot. The above sum corresponds to the exponential 
map. We do not have enough space in this paper to develop the theory in a more 
rigorous way. The reader can easily state that ``following a 
constant velocity'' makes sense for the four types of joints defined in 
Table~\ref{tab:joints}. We refer to~\cite{riemanian-optim2008} Chapter~5 for 
details about Riemanian geometry.

@@

We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision. This 
corresponds to the unconstrained function \texttt{computeIterate} in 
Algorithm~\ref{algo:gradient}.
When a collision is detected, we introduce a constraint and perform a new 
iteration from $\xx_i$ as explained in the next section. We use a continuous 
collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
paths and to return the first colliding configuration along a path. This step corresponds to \texttt{validatePath} in Algorithm~\ref{algo:gradient}.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=9cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The corresponding constraint will be computed in configuration $\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

@@

Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. Let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free (see 
Figure~\ref{optim_grad}). Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, we will only consider the first case. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

@@

Let $M_1 \in SE(3)$ (resp. $M_2 \in SE(3)$) denote the homogeneous transformation 
between frames of $\body_1$ (resp. $\body_2$) and the global frame in configuration 
$\xx_i(\tcolli)$. Let $\mathbf{u}$ be the unit vector 
linking the contact points in configuration 
$\xx_{i}(t_{coll\ i})$ (the latest collision-free path), expressed in the local 
frame of $\body_1$:
$$
\mathbf{u} = \frac{M_{1}^{-1} (\xx_i) M_2 (\xx_i) \po_2 - \po_1}{\|M_{1}^{-1} 
(\xx_i) M_2 (\xx_i) \po_2 - \po_1\|}
$$

Notice that $\mathbf{u}$ is well defined since configuration 
$\xx_{i}(t_{coll\ i})$ is collision-free.
Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
$$
f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
$$
The constraint is defined by
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
As this constraint is non-linear, it is linearized around $\xx_{i}$.

Let $q_{k\,i}$ denote the waypoint $k$ along path $x_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $q_{k\,i}$, $q_{k+1\,i}$, with $0\leq k\leq n$. There exists 
$\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
$$
The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
Algorithm~\ref{algo:gradient}.
Then, a line is added in the contraint Jacobian matrix (function 
\texttt{addCollisionConstraint}), and the constraint 
Jacobian becomes:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}

@@

Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP) (function \texttt{computeIterate} in 
Algorithm~\ref{algo:gradient}).

@@

\subsection*{Convergence Analysis}
For iteratively constrained problem, the risk of falling in an infinite loop 
that keeps adding a redundant constraint exists.
In this section, we demonstrate that, given the constraint definition and the 
function \texttt{solveRedundantConstraint}, our 
algorithm will converge.
Under the previous notations, it is straight-forward that
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
Let us assume that
\begin{equation}\label{eq:asumption-1}
L_{k+1}\p_i \not= 0.
\end{equation}
We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then
$$
\kernel J_i \not= \kernel J_{i+1}.
$$
From~(\ref{eq:decreasing-kernel}), we deduce that
$$
\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
$$
This results proves that each added constraint is linearly independant from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

\subsubsection*{About asumption~(\ref{eq:asumption-1})}%
Let us consider the following expression:
$$
\frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
This expression represents the velocity of point $\po_2$ in the reference frame of Body~1 projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
\begin{equation}\label{eq:trajectory}
\conf (t) = \xx_{i}(t_{coll\ i}) + t\, \alpha_i \p_{i}.
\end{equation}
The motion is defined by taking a constant abscissa $t_{coll\ i}$ and be moving the 
path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
reached for $t=1$.
This expression being equal to 0 requires either
\begin {enumerate}
\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
\item the velocity of $\po_2$ in the reference frame of Body~1 is orthogonal to 
$\mathbf{u}$ at contact ($t=1$).
\end {enumerate}
The first case is very unlikely and almost never encountered in real applications, 
except if the user builds such a case on purpose.
The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$. In this 
case, we do not directly insert the new constraint. Instead, we process intermediate 
iterations until the redundancy vanishes, as explained in function 
\texttt{solveRedundantConstraint}.

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
\State $solved \gets false$
\State $minReached \gets false$
\While{($\textbf{not}(solved)$)}
	\State $\p = \texttt{computeIterate}()$
	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx \gets \xx_{Free} + \frac{\alpha}{2}\,\p$
	\If{(\texttt{validatePath}($\xx_1$))}
		\State $\xx_{Free} \gets \xx$
	\Else
		\State $\xx_{Coll} \gets \xx$
	\EndIf
	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
\EndWhile
\end{algorithmic}
\caption{\texttt{solveRedundantConstraint}()} \label{algo:solveRedundant}
\end{algorithm}

@@

Doing so, either the new constraint makes the dimension of the kernel decrease, or 
makes the configuration in collision ($\xx_{i+1}$ or $\xx_{i+2}$ depending on the 
case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
closer and closer. And the continuity of $f$ implies that
$$
\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
body~1 along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
contact. We already handled this case previously, admitting that the velocity norm 
would not vanish. Therefore, is~\ref{eq:asumption-1} verified.

@@

In this part we describe the GB path-optimizer Algorithm~\ref{algo:gradient} 
according to the previous step definitions. The main difficulty here is to handle 
the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
be traveled through, and also to return a collision-free path. Typically, we 
chose $\alpha_{init} = 0.2$ or 0.1 to process small steps. $\alpha_{init} = 0.5$ is 
more encountered in the optimization literature, procuring larger steps but with 
more risks of collision.

>>>>

% The following packages can be found on http:\\www.ctan.org
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{algorithm}
\usepackage{algpseudocode}  %\usepackage[noend]{algpseudocode}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{proof}
\newtheorem{property}{Property}
\usepackage[numbers]{natbib}
\algnewcommand\algorithmicinput{\textbf{input:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\algorithmicoutput{\textbf{output:}}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
\makeatletter
\makeatother
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!80!black},
    citecolor={blue!80!black},
    urlcolor={blue!80!black}
}

@@

\begin{abstract}
Most algorithms in probabilistic sampling-based path planning compute 
collision-free paths made of straight line segments lying in the configuration 
space. Due
to the randomness of sampling, the paths make detours that need to be optimized.
The contribution of this paper is to propose a gradient-based (GB) algorithm that
transforms a polygonal collision-free path into a shorter one, while both:
\begin{itemize}
\item requiring only collision checking, and not any time-consuming obstacle distance computation nor geometry simplification,
\item constraining only part of the configuration variables that may cause a 
collision, and not the entire configurations, and
\item reducing parasite motions that are not useful for the problem resolution without any assumption.
\end{itemize}
The algorithm is simple and requires few parameter tuning. Experimental
results include navigation and manipulation tasks, e.g. an robotic arm 
manipulating throw in boxes and a PR2 robot working in a kitchen environment. 
Comparisons with a random shortcut optimizer and a partial shortcut are also provided.

@@

Motion planning for systems in cluttered environments has been addressed for more
than thirty years~\citep{ref-motionplan},~\citep{ref-pianomover}. Most planners
randomly sample the system configuration space~\cite{ref-spaceconfig} in
order to find a collision-free path, linking the initial and final configurations. 
The main issue using these techniques is that the computed path makes unnecessary 
detours and needs to be post-processed before being executed by a virtual or real 
robot. Alternative strategies exist however.
\begin{itemize}
\item Planning by path-optimization~\cite{itomp2012},~\cite{voronoiOMP} where
obstacle avoidance is handled by constraints or cost using computation of the
nearest obstacle distance. Most of these planners are using non-linear
optimization~\cite{BettsNonlinopt} under constraints. Such planners provide close-to-optimality paths and have smaller time computation
for easy problems, but they are mostly unable to solve narrow passage issues.

@@

For this shortening purpose, random shortcut (RS) methods are
still very 
popular~\cite{Sekhavat-Svestka1998,Geraerts04clearancebased,HauserFastSmooth}. 
However, RS requires fine 
tuning of the termination condition (see 
(Algorithm~\ref{algo:random-shortcut})) and is no efficient due to randomness for long 
trajectories where only a small part needs to be optimized. 
(Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS will 
always fail to optimize the initial path, since it cannot decouple the 
robot degrees of freedom (DOF) on which the optimization occurs, contrary to our 
algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} 
that shortcut only one or a group of DOF.

\begin{figure}[H]
	\centering
	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
	\caption{Example of a path which random shortcut will never manage to 
	optimize: each shortcut tentative will provide a collision or will not 
	decrease the path length. The optimal path belongs to the $x-y$ plane 
	containing $\conf_{init}$ and $\conf_{final}$.}
	\label{decoupled_DOF_optimization}
\end{figure}

@@

\section{RELATED WORK}
Previous work on path optimization has been conducted. CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
input. It minimizes a time invariant cost function using efficient covariant
hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
high velocities) and an obstacle avoidance term, provided by the distance to the 
nearest obstacle for each iteration of the trajectory. Calculating these nearest 
distances however is time-consuming because the distances between all pairs of 
objects must be computed at each time step along the path. To reduce the 
computation time, the method starts by building offline a map of distances that 
will be called during the optimization at the requested time. Besides, meshes 
are pre-processed into bounding spheres so that distances are computed faster 
at the cost of a geometry approximation.

@@

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints. 
~\cite{Guernane2011}, also based on random shooting, is guiding configuration-
generation with local holonimic considerations. But this method remains only 
locally optimal, and is not addressing high-DOF problems.
~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
using medial axis retractation for clearance and a PRS (applied only on random DOF) 
which can address the problem of (Figure~\ref{decoupled_DOF_optimization}) but is 
relatively slower than RS. Moreover, PRS is not using the information returned by 
the checker of which links are colliding to guide which DOF or group of DOF are 
relevant to shortcut, contrary to our method.

In some way, our method shares similarities with~\cite{PanSmoothSplineShort} 
since this latter work relies on collision checking and backtracks when an 
iteration is detected in collision, instead of trying to constantly satisfy 
distance constraints. Unlike our method however, anti-collision constraints are 
handled by interpolating configurations which, at some points of the trajectory, 
freeze the whole robot configuration instead of a relevant subpart. Thus, this 
method cannot solve the (Figure~\ref{decoupled_DOF_optimization}) problem.

@@

A robot is defined by a kinematic chain composed of a tree of joints. We denote 
by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $1\leq i\leq 
m$ is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
is the dimension of $J_i$ in the configuration space, to the space 
of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
in the frame of its parent. In the examples shown in this paper, we consider 4 
types of joints described in (Table~\ref{tab:joints}).
\begin{table}[b]
\centerline {
  \begin{tabular}{cccc}
    Name & dimension & config space & velocity\\
    \hline
    translation & 1 & $\real$ & $\real$\\
    bounded rotation & 1& $\real$ & $\real$\\
    unbounded rotation & 2 & $\Sone\subset\real^2$ & $\real$\\
    $SO(3)$ & 4 & $\Sthree\subset\real^4$ & $\real^3$
  \end{tabular}
}
\caption {Translation and rotation joint position are defined by 1 
parameter corresponding respectively to the translation along an axis and a 
rotation angle around an axis. Unbounded rotation is defined by a point on the 
unit circle of the plane: 2 parameters corresponding to the cosine and the 
sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
velocity of translation and unbounded rotation joints is the derivative of the 
configuration variable. The velocity of an unbounded rotation joint corresponds 
to the angular velocity. The velocity of a $SO(3)$ joint is defined by the 
angular velocity vector $\omega\in\real^3$.}
\label{tab:joints}
\end{table}
A configuration of the robot
$$\conf = (\underbrace{q_1,\cdots,q_{n_1}}_{J_1},\underbrace{q_{n_1+1},
\cdots,q_{n_1+n_2}}_{J_2},\cdots q_n),\ n\triangleq\sum_{i=1}^m n_i$$
is defined by the concatenation of the joint configurations. The configuration 
space of the robot is denoted by $\CS\subset\real^n$.

@@

Note that the definitions above stem from the Riemanian structure of the 
configuration space of the robot. The above sum corresponds to the exponential 
map. We do not have enough space in this paper to develop the theory in a more 
rigorous way. The reader can easily state that ``following a 
constant velocity'' makes sense for the four types of joints defined in 
(Table~\ref{tab:joints}). We refer to~\cite{riemanian-optim2008} Chapter~5 for 
details about Riemanian geometry.

@@

We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision. This 
corresponds to the unconstrained function \texttt{computeIterate} in 
(Algorithm~\ref{algo:gradient}).
When a collision is detected, we introduce a constraint and perform a new 
iteration from $\xx_i$ as explained in the next section. We use a continuous 
collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
paths and to return the first colliding configuration along a path. This step corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}).

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=9cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The 
	corresponding constraint will be computed in the backtracked configuration 
	$\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

@@

Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free. Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, we will only consider the first case. Reasoning 
about the second case is similar, except that the 
constraint is on the position of $\body_1$ with respect to the environment.

@@

Let $M_1 \in SE(3)$ (resp. $M_2 \in SE(3)$) denote the homogeneous transformation 
between frames of $\body_1$ (resp. $\body_2$) and the global frame in configuration 
$\xx_i(\tcolli)$. Let $\mathbf{u}$ be the unit vector 
linking the contact points $\po_1$ and 
$\po_2$ in configuration $\xx_{i}(t_{coll\ i})$, $\xx_{i}$ being the latest collision-free path:
$$
\mathbf{u} = \frac{M_{1}^{-1} (\xx_i) M_2 (\xx_i) \po_2 - \po_1}{\|M_{1}^{-1} 
(\xx_i) M_2 (\xx_i) \po_2 - \po_1\|}
$$

Notice that $\mathbf{u}$ is well defined since configuration 
$\xx_{i}(t_{coll\ i})$ is collision-free.
Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
$$
f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
$$
The constraint is defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
As this constraint is non-linear, it is linearized around $\xx_{i}$. Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
The constraint on the position at $t_{coll\,i}$ is a function that depends only
on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq n$. There exists 
$\beta\in[0,1]$ such that for any $j$:
$$
\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
$$
The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
(Algorithm~\ref{algo:gradient}).
Then, a line is added in the contraint Jacobian matrix (function 
\texttt{addCollisionConstraint}), and the constraint 
Jacobian becomes:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}

@@

Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP) (function \texttt{computeIterate} in 
(Algorithm~\ref{algo:gradient})).

@@

\subsection*{Convergence Analysis}
For iteratively constrained problem, the risk of falling in an infinite loop 
that keeps adding a redundant constraint exists.
In this section, we demonstrate that, given the constraint definition and the 
function \texttt{solveRedundantConstraint}, our 
algorithm will converge.
Under the previous notations, it is straight-forward that
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
Let us assume that:
\begin{equation}\label{eq:asumption-1}
L_{k+1}\p_i \not= 0.
\end{equation}
We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
$$
\kernel J_i \not= \kernel J_{i+1}.
$$
From~(\ref{eq:decreasing-kernel}), we deduce that:
$$
\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
$$
This results proves that each added constraint is linearly independant from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

\subsubsection*{About asumption~(\ref{eq:asumption-1})}%
Let us consider the following expression:
$$
\frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
This expression represents the velocity of point $\po_2$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
\begin{equation}\label{eq:trajectory}
\conf (t) = \xx_{i}(t_{coll\ i}) + t\, \alpha_i \p_{i}.
\end{equation}
The motion is defined by taking a constant abscissa $t_{coll\ i}$ and be moving the 
path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
reached for $t=1$.
This expression being equal to zero requires either
\begin {enumerate}
\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
\item the velocity of $\po_2$ in the reference frame of $\body_1$ is orthogonal to 
$\mathbf{u}$ at contact ($t=1$).
\end {enumerate}
The first case is very unlikely and almost never encountered in real applications, 
except if the user builds such a case on purpose.
The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$. In this 
case, we do not directly insert the new constraint. Instead, we process intermediate 
iterations until the redundancy vanishes, as explained in function 
\texttt{solveRedundantConstraint}.

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
\State $solved \gets false$
\State $minReached \gets false$
\While{($\textbf{not}(solved)$)}
	\State $\p = \texttt{computeIterate}()$
	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx \gets \xx_{Free} + 0.5\alpha\,\p$
	\If{(\texttt{validatePath}($\xx_1$))}
		\State $\xx_{Free} \gets \xx$
	\Else
		\State $\xx_{Coll} \gets \xx$
	\EndIf
	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
\EndWhile
\end{algorithmic}
\caption{\texttt{solveRedundantConstraint}()} \label{algo:solveRedundant}
\end{algorithm}

@@

Doing so, either the new constraint makes the dimension of the kernel decrease, or 
makes the configuration in collision ($\xx_{i+1}$ or $\xx_{i+2}$ depending on the 
case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
closer and closer. And the continuity of $f$ implies that
$$
\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
$\body_1$ along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
contact. We already handled this case previously, admitting that the velocity norm 
would not vanish. Therefore, Equation~(\ref{eq:asumption-1}) is verified.

@@

In this part we describe the GB path-optimizer (Algorithm~\ref{algo:gradient})
according to the previous step definitions. The main difficulty here is to handle 
the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
be traveled through. Typically, we 
chose $\alpha_{init} = 0.2$ or 0.1 to process small steps. $\alpha_{init} = 0.5$ is 
more encountered in the optimization literature, procuring larger steps but with 
more risks of collision.
